### 2013-07-15 LATEX/beg.draft.2013.tex
Included discussion of results and cleaned up figures

\documentclass[11pt]{article}
\usepackage[dvips]{graphics}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{chicago}
\usepackage{rotating}
\usepackage{hyperref}
\usepackage{sectsty}

\addtolength{\topmargin}{-0.375in}
\addtolength{\oddsidemargin}{-0.35in}
\addtolength{\evensidemargin}{-0.35in}
\addtolength{\textheight}{0.5in}
\addtolength{\textwidth}{0.5in}
\addtolength{\parskip}{0.05in}
\renewcommand{\baselinestretch}{1.5}
\newcommand{\vsp}{\vspace{\baselineskip}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\gr}[1]{\ensuremath  {\frac{\dot {#1}}{#1}}}

\begin{document}
\renewcommand{\baselinestretch}{1}
\author{{\Large Valentin Bolotnyy} \\
Federal Reserve Board \and
{\Large Rochelle M. Edge} \\
Federal Reserve Board \and
{\Large Luca Guerrieri}\\ Federal Reserve Board} 
\title{Stressing Bank Profitability for Interest Rate Risk$^*$ \\ \vspace{0.5cm} Preliminary and Incomplete \vspace{0.5cm}}

\date{\today}
\maketitle

\begin{abstract}
Regular bank stress testing is one of the three complementary reforms that has been made to U.S. capital regulation in response to the shortcomings of the pre-crisis regime.
The form of stress testing mandated by the Dodd-Frank Act for the Federal Reserve to undertake is one based around a small number of scenarios.  This approach involves specifying macroeconomic and financial scenarios that represents stressful conditions over the time horizon of the stress test and then projecting, for each bank holding company in the stress test, its losses, income, and path of capital \textit{given that scenario}. In this paper, we focus on the path of income. In particular, we consider the ability of empirical models to translate developments in interest rates into bank net interest margins (NIMs).
\end{abstract}

\vspace{.125in}

\noindent Key Words: Net Interest Margins, Interest Rate Risk, Forecasting, Stress Tests.

\vspace{.125in}

\noindent JEL Classification: C53, E47, G21

\vspace{1.5cm} \renewcommand{\baselinestretch}{1} \normalsize
\noindent $^*$ The views
expressed in this paper are solely the responsibility of the
authors and should not be interpreted as reflecting the views of
the Board of Governors of the Federal Reserve System or of any
other person associated with the Federal Reserve System.
Contact Information: valentin.bolotnyy@frb.gov; rochelle.m.edge@frb.gov; luca.guerrieri@frb.gov.

\renewcommand{\baselinestretch}{1.5} \normalsize
\thispagestyle{empty}

\newpage
\setcounter{page}{1}

\section{Introduction}

\vspace{-0.1in}
Regular bank stress testing (and the capital planning that it implies) is one of the three complementary reforms that has been made to U.S. capital regulation in response to the shortcomings of the pre-crisis regime.\renewcommand{\baselinestretch}{1}\footnote{See Governor Tarullo's November 2011 speech ``The Evolution of Capital Regulation'' for this characterization of stress testing. The two other reforms that he mentions are improvements to the traditional, individual firm approach to capital regulation and the introduction of a macroprudential component to capital regulation.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} While stress testing can take many forms---for example, being based on a small number of scenarios, being based on a large number of scenarios, or even being undertaken in reverse form to achieve a certain adverse outcome---the form of the stress tests mandated under the Dodd-Frank Act for the Federal Reserve to undertake and the form being used by the Federal Reserve for its annual Comprehensive Capital Analysis and Review (CCAR) is one based around a small number of scenarios.  This approach involves specifying a small number of macroeconomic and financial scenarios that represent stressful conditions over the time horizon of the stress test and then projecting, for each bank holding company (BHC) in the stress test, its losses, income, and path of capital \textit{given that scenario}. Accordingly, there are at least two important parts to conducting bank stress tests (i)~formulating macroeconomic and financial scenarios that should be stressful to banks and (ii)~developing econometric models that are able to translate such stressful conditions into losses, income, and capital.

We examine how well a range of time-series models can explain NIMs. We emphasize the models' pseudo-out-of-sample \textit{conditional} forecast performance. We adopt this pseudo-out-of-sample approach for the simple reason that the evaluation of models based on their in-sample forecast performance, may lead to in-sample over-fitting and poor out-of-sample performance. An explicit focus on out-of-sample performance makes forecast uncertainty explicit.
Moreover, concentrating on conditional forecasts, allows us to separate the task of modeling and forecasting NIMs from that of forecasting interest rates, which is itself a sizable topic (see Duffee, 2012).

When developing econometric models to translate stressful economic conditions into key bank financial-statement variables, it is important to consider what types of developments are likely to be stressful to banks and what particular economic variables are likely to capture these stresses. It is then critical that the models being used for stress testing be able to translate developments in these economic variables into conditional forecasts for key bank financial-statement variables.  In this paper, we examine this property for models of bank net interest margins (NIMs), defined as the ratio of net interest income (NII) to interest earning assets, where NII is the difference between a bank's interest income and interest expenses.

In determining which variables to include in our model of NIMs we look to the variables that have been emphasized by two existing strands of the literature on NIMs.  The first strand emphasizes the link between interest rates of different maturities and NIMs and is more in the macro-banking tradition. The second literature focuses on the optimal margin set by banks based on considerations like the competitive structure of the lending and deposit markets in which the bank participates, the degree of managerial risk aversion, and the degree of uncertainty associated with the future path of interest rates and is much more in the micro-banking tradition.

Our motivation for focusing on models of NIMs is fourfold. First, losses from interest rate risk have been a central feature of banking crises. Second, net interest rate margins capture income from the maturity transformation services at the center of the traditional banking model. This source of income has been lackluster given the current configuration of interest rates. Third, modelling of NIMs has received scant attention so far. Fourth, the ability of producing reliable conditional forecasts for a key component of net income is at the core of the scenario-based stress testing of banks. We elaborate on each of these four parts of our motivation below.

Bank losses that arise from banks' having to pay out more in interest on their liabilities than they are earning from their assets should not be overlooked as an important source of risk to the banking and broader financial sector.  To be sure, losses that arise from this source of risk---called, interest rate risk---were much less prominent in the most recent crisis relative to bank losses that stem from loan defaults---that is, credit risk---but there are ample examples of other earlier banking crises in which adverse developments in NIMs have played a significant role.  The most familiar of these from the U.S. perspective is the Savings and Loans (S\&L) crisis that started in the early 1980s with short-term interest rates rising above long-term interest rates---in part, as a result of the Volcker disinflation.  This resulted in interest expenses rising above interest income, NII and NIMs turning negative (in the thrift sector), and sizable losses and ultimately a substantial number of bank failures.

Likewise, the Nordic banking crises of the late 1980s and early 1990s (for Finland, Sweden, and Norway)---which represent three of Reinhart's and Rogoff's~(2009) ``big five'' banking crises---were associated with elevated short-term interest rates that depressed NII and NIMs and lead to sizable bank losses and failures. In the Nordic crises, high short-term interest rates resulted from these countries attempting to maintain currency pegs with the Deutschemark in the face of the tight German monetary policy that accompanied its loose fiscal policy associated with re-unification spending. Depressed NIMs were also at the center of the U.K.'s secondary banking crisis of the early 1970s. A sudden increase in short-term interest rates abruptly drove up banks' costs of funds,  leading to extensive bank losses and failures.  To be sure, large increases in short-term rates and declines in net interest income were not the \textit{only} reason for any of these crises---indeed, there was at least one major class of asset prices that collapsed in each of these crises---but short-term interest rates---and, in particular, their elevated levels relative to long-term interest rates---were significant and in most cases equivalent in importance to the other causes of the crisis.

The recent pace of bank profit growth has been lackluster and is becoming an increasing source of concern to market commentators and regulators alike. In recent years, bank net income growth has been largely supported by unsustainable cost cutting and reductions in provisioning, rather than from sources like NII.  NII growth, for example, (in both nominal and real terms) has been well below average in recent years while NIMs (NII relative to interest earning assets) have also been depressed. Given their already low level, any adverse development that puts further downward pressure on NIMs could be very stressful for banks and, as such, it is important that the economic developments (likely involving interest rates) that would drive these outcomes are well-modeled.

Despite the importance of NIMs for the financial condition of banks and despite the attention that it receives from bank equity analysts, the modeling of NIMs has received much less attention in the risk-management literature relative to the modeling of bank variables related to credit risk.  To be sure, the recent increase in the importance of stress testing has led to the development of a number of models that link NIMs to macroeconomic variables---such as, short-term Treasury rates and the slope of the Treasury yield curve---with Kovner \textit{et al}.~(2011) and Covas \textit{et al}.~(2012) being two notable examples for the U.S.  However, this research places relatively less emphasis on the conditional forecasting properties of these models, which is important given that these models are ultimately to be used to project bank losses, income, and capital given macroeconomic scenarios.\renewcommand{\baselinestretch}{1}\footnote{Covas \textit{et al}. do present some forecast evaluation results in order to demonstrate the benefits of focusing on quantile projections. However, they only report results for net chargeoffs and pre-provisioning net revenue and their focus is on density forecast evaluation.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}

Scenario based stress tests are predicated on the ability to produce reliable conditional forecasts for key measures of profitability such as net interest margins. This kind of stress test program is not limited to the U.S. banking system. Stress testing of banks based on scenarios was undertaken during the Scandinavian banking crisis to shore up confidence in the ability of banks to withstand the crisis. Most extensively, the European Banking Authority conducted stress tests based on scenario analysis throughout the European Union in 2009, 2010, and 2011. The results of such programs can only be effective in restoring confidence from the public in the midst of a crisis to the extent that the public can be convinced of the ability of regulators to assess the consequences of their scenarios on banks. If forecast methods at the frontier of efficiency have trouble disentangling the implications of baseline and severe stress scenarios, the ability of stress test programs to avoid panic is called in question.

%In summing up our motivation on the importance of modeling NIMs and studying interest rate risk, we would also note that the Federal Reserve chose to feature interest rate risk in one of the scenarios specified for the 2013 %Dodd-Frank Act (DFA) stress tests.  In particular, for the 2013 DFA stress tests, the Federal Reserve specified the adverse scenario as one in which the Treasury yield curve shifted up and flattened.  As noted by the Federal %Reserve in its (proposed) Policy Statement on scenario design, the Board varies ``the approach it uses for the adverse scenario each year so that the results of the scenario provide the most value to supervisors, in light of %current condition of the economy and the financial services industry.''

We proceed as follows. Section~2 provides a review of the literature on NIMs with an eye to cataloguing variables that have been found useful to explain the evolution of NIMs. Section~3 turns to the models that we develop to forecast NIMS. Section~4 describes the estimation data and Section~5 the estimation results. Section~6 concludes.

\section{Related literature and variable selection}

\vspace{-0.1in}
There are two strands in the literature on modeling NIMs.  The first strand emphasizes the link between interest rates of different maturities and NIMs and lies more in the macro-banking tradition.  The second strand focuses on the optimal margin set by banks and lies more in the micro-banking tradition.

Kovner \textit{et al}. and Covas \textit{et al.} are in the macro-banking tradition. Both link NIMs to the slope of the yield curve and a short-term interest rate.  Banks' earnings from interest income reflect two key services that banks supply---maturity transformation services (that is, borrowing short and lending long) and deposit transactions services---and the returns to these services are reflected by these two interest rate variables.  The slope of the yield curve reflects the return to banks from maturity transformation and thus is expected to enter NIM models with a positive sign. The short-term market rates in NIM models reflect the fact that bank deposit rates, while typically lower (since they provide transactions services), are constrained by the zero lower bound.  As such, the short rate places an upper limit on what banks can earn from the provision of their transactions services. As a result, the short rate is expected to enter the model, also, with a positive sign. Figure \ref{figure_nims_components} highlights the covariation of NIMs with the 3-month Treasury yield.  In the figure NIMs are split into income and expenses components, both normalized by interest earning assets. In this and in the other figures described in this section, we focus on average for the top 25 bank holding companies in the United States by assets, assessed quarterly.  The evolution of both components of NIMs bracket the evolution of the three month Treasury rate and the contours are strikingly similar. Figure \ref{figure_nims_rates} shows the evolution of NIMs against that of the 3-month and the 10-year Treasury yields.

Both Kovner \textit{et al}. and Covas \textit{et al}. measure the slope of the yield curve in their respective NIM models with the 10-year to 3-month Treasury term spread and short-term market rates with the 3-month Treasury bill rate, as do English~(2002) and English \textit{et al}.~(2012)---two other papers in this tradition.  All of these papers estimate NIM models with slightly different purposes in mind.  The NIM models estimated by Kovner \textit{et al}. and Covas \textit{et al}. are developed for stress testing and are part of much larger aggregative models of a number of components of banks' financial statements. These models are both estimated on panel data with the most notable difference between these models being the estimation techniques they employ: Least squares regression in the case of Kovner \textit{et al}. and quantile regression in the case of Covas \textit{et al}. Figure \ref{figure_nims_factors} shows the evolution of average NIMs against ``observed'' level, slope, and curvature factors.\footnote{Following Diebold and Li (2006), we compute the observed factors, respectively as: the average of the 3-month, 2-year, and 10-year Treasury yields; the difference between the 10-year and the 3-month yields; and 10-year yield minus two times the 2-year yield plus the 3 month yield.}. As made explicit by the figure, NIMs comove strongly with the slope factor over the entire sample, but the level and curvature factors also move closely with NIMs over certain subperiods.

Like our paper, English focuses exclusively on NIMs, although he does so for a number of countries (that is, all G-7 countries except for France, and for Australia, Norway, Sweden, and Switzerland).  Addditionally (and as we also do later), he considers asset yields and liability yields separately---the two measures that when differenced imply NIMs.  English's main focus is understanding how well the aggregate banking sectors in the countries he studies appear to have managed the interest rate risk that they face on their earnings, as such he uses aggregate data (rather than firm-level data) and does not focus on the conditional forecasting performance of his model.

Finally, English \textit{et al} consider the relationship between NIMs and the slope of the yield curve and short-term rates for a panel of banks, however, in contrast to the other studies they allow for the specific coefficients estimated for each bank to vary as linear functions of bank balance sheet ratios that one \textit{a priori} might think would be relevant to net interest income.  In particular, these ratios include the maturity gap between bank assets and liabilities, the deposit share of liabilities, and the loan share of assets. Again, since English \textit{et al} are focussed on understanding the determinants of NIMs they do not consider conditional forecast performance.

A separate strand of the NIM modeling literature comes from the micro-banking tradition, which focuses on the determinants of the loan rates and deposit rates that banks set as implied by the maximization of their profits. Papers in this tradition---such as, Ho and Saunders~(1981), Angbazo~(1997), and Saunders and Schumacher~(2000)---are based around what is known as the ``dealership'' model of banks; that is, banks in these models are viewed as dealers in the credit market, acting as an intermediary between demanders and suppliers of funds.  Banks are modeled as being able to raise funds from and lend funds to short-term money markets as well as being able to raise deposits from and make loans to nonfinancial agents.

In borrowing or lending in money markets banks face a market rate, which can vary over time, while in raising deposits from or extending loans to nonfinancial agents banks have the ability to set rates.  In raising deposits, banks can pay a below market rate that is fixed for the full life of the deposit and in extending loans they can earn an above market rate that is also fixed for the full maturity of the loan.  Importantly, the life of the deposit is the same as that of the loan, reflecting the fact that while core deposits can, in principal, be withdrawn on demand from a bank they are in practice extremely sticky.  Interest rate risk arises in this model from the fact that deposit-taking and lending opportunities do not arrive at the bank at precisely the same time but rather arrive at any time with some probability, where these probabilities---which are different for deposits and lending opportunities---depend on the interest rates set by the bank on deposits and loans.  Consequently, banks can find themselves holding deposits without loans, that they must then invest at variable market rates (and that could adversely fall below the deposit rate), or loans without deposits, that they must then fund at variable market rates (that could adverse rise above the lending rate).

The solution to the bank's profit-maximization problem is the spread between the loan rate and the deposit rate and this has two parts. The first part is the spread that a bank would set if it were completely risk neutral. This depends only on the degree of competition faced by the bank in the loan markets and the deposit markets in which it operates, where, in the model, this is captured by the slope of the loan demand schedule and slope of the deposit supply schedule facing the bank. The second part of the spread reflects banks' risk aversion to the situation of either having deposit supply without corresponding loan demand or having loan demand without the corresponding deposit supply. Accordingly, the second part of the spread is the compensation for this risk that banks require in order for them to be willing to take deposits and lend. This depends positively on the banks' degree of risk aversion as well as the volatility of interest rates.

In Ho and Saunders the degree of competition faced by the bank in loan and deposit markets does not enter their estimated model, since they argue that this variable is quite slow moving and their analysis is undertaken only over a period of three years.  Saunders and Schumacher do consider the degree of competition in their analysis--which is cross-country in nature---although this takes the form of a country dummy variable in the regressions. In all cases the banks' degree of risk aversion does not enter the specification for estimated models of NIMs since this is difficult to measure, although the volatility of interest rates does and this is measured by the annual or quarterly standard deviation of weekly Treasury rates of different maturities (\textit{e.g.}, 3-month, 1-year, 2-year, 3-year, and 5-year rates). Figure \ref{figure_nims_competition} shows the evolution of NIMs against the evolution of the measure of competition that we consider -- the shadow banking sector share of assets in the overall banking sector comprising both banks and shadow banks.  For most of the sample this share increases, but the decline is arrested and partially reversed towards the end of the sample. [TO DO: add a panel to that figure that includes a measure of volatility].

Most papers in this literature also add a few more controls to the model before estimation.  These variables include banks' payments of implicit (instead of, explicit) interest on deposits;\renewcommand{\baselinestretch}{1}\footnote{This takes place through service charge remissions and other types of depositor subsidies due to regulatory restrictions of explicit interest payments and is measured by the difference between noninterest expenses and noninterest revenue, all divided by interest earning assets.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} banks' non-interest-bearing reserve requirements;\renewcommand{\baselinestretch}{1}\footnote{The difference between noninterest expenses and noninterest revenue captures the influence of the economic cost of funds over and above the published interest expense and is measured by non-interest bearing reserves divided by total assets.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} banks' default risk, since this should imply an additional premium charged on the rate of the banks loans given the potential loss in bank capital;\renewcommand{\baselinestretch}{1}\footnote{This is measured by net chargeoffs divided by total loans.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} and, bank capital, since capital is costly and banks with higher capital may seek to cover this with higher lending rates.  Note that all of the empirical analysis in the ``dealership'' model literature for studying NIMs is undertaken at the firm-level and as such many of the \textit{ad hoc} controls that are added to the model are included so as to address cross-firm differences.

We would note that Angbazo does add default risk to the bank's profit-maximization problem and, as such, default risk does enter his model of NIMs in a more rigorous way.  Here he finds that for money center banks net interest margins are more sensitive to default risk than they are to interest-rate risk, while super-regional and regional banks are sensitive to interest-rate risk but less so to default risk. These findings, he notes, are consistent with the money center banks' greater concentration of short-term assets and greater use of off-balance sheet (OBS) hedging instruments.

There is a further strand of the micro-banking literature looking at determinants of banks' NIMs---which includes papers by Zarruk~(1989) and Wong~(1997)---that we mention but do not discuss at length.  In these models deposits and lending opportunities arrive at the same time and depend on the deposit and loan rates being set by banks but there is some stochastic element to the volume of deposits that banks will raise given the interest rate they set, thus requiring the banks in the model, who are risk average, to borrow or lend in the federal funds market.  These models also include features like administrative costs on deposits and loans as well as costs associated with deposit insurance premia and the primary use of these models is for comparative static exercises that examine how changes in these parameters---as well as considerations like uncertainty---will affect bank spreads.  Ultimately, however, these papers are not empirical and since the majority of their theoretical findings confirm those of those of papers like Ho and Saunders, Angbazo, and Saunders and Schumacher, we do not delve too deeply into them here.

In summing up, in addition, to short- and long-term interest rates, a few variables stand out as relevant for our analysis.  The degree of competition faced by the bank in the loan markets and the deposit markets in which it operates stands out as an important consideration, especially given that the sample we use in our analysis is much longer than the samples considered by Ho and Saunders, Angbazo, and Saunders and Schumacher.  Interest rate volatility, also appears important, since this has likely also changed over our sample.  The other variables that previous authors in the micro-banking tradition have added to their models in an \textit{ad hoc} way we hold-off for the time being from adding to our analysis.  This is largely because our analysis at this stage is at the aggregate level and these additional variables appear to have been added to explain variations in NIMs across firms.

\section{Models for forecasting NIMs}

We construct forecasts from a wide array of standard methods taken from the forecasting literature. We use data with quarterly frequency and consider a forecast horizon up to ten quarters long. We are interested in forecasts for NIMs conditional on the variables in scenario $I$, which we denote $NIM_{t+s/I}$. Here are the methods considered:

\vspace{-0.1in}
\noindent \textbf{1. Equal-weighted forecast combination.} We regress NIMs on each yield $r(\tau )_{t-1}$ separately: \
\[
NIM_{t}=c_{\tau }+\rho _{\tau }NIM_{t-1}+\sum_{j=1}^4\gamma _{\tau,j}r(\tau)_{t-j}+{\eta
_{\tau ,t}}
\]
Each regression is used to produce an $s$ step-ahead forecast of NIMs conditional on yields on treasuries $r(\tau)$ in scenario $I$ with maturity $\tau$ observed through period $t+s-1.$ This conditional forecast is denoted by $NIM_{\tau,t+s/I}.$ The equal-weighted forecast combination is then given by:
\[
NIM_{t+s/I}=\sum_{\tau }\frac{NIM_{\tau ,t+s/I}}{N},
\]
where $N$ is the number of maturities considered.\footnote{Experimentation with alternative lag structures led to no improvement in the forecast results reported below.}

\noindent \textbf{2. A dynamic factor model for NIMs.} We include NIMs in the dynamic factor model for the yield curve developed by Diebold and Li (2006). Using the state space representation, the state equations of our dynamic factor model is given by:
\[
\left(
\begin{array}{c}
{L_{t+1}-\mu _{L}} \\
{S_{t+1}-\mu _{S}} \\
{C_{t+1}-\mu _{C}} \\
\widehat{NIM}_{t+1}-\mu _{\widehat{NIM}}%
\end{array}%
\right) {\normalsize \ =\ }\left(
\begin{array}{cccc}
a_{11} & a_{12} & a_{13} & 0 \\
a_{21} & a_{22} & a_{23} & 0 \\
a_{31} & a_{32} & a_{33} & 0 \\
a_{41} & a_{42} & a_{43} & a_{44}%
\end{array}%
\right) {\normalsize \ }\left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}} \\
\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
\end{array}%
\right) {\normalsize +}\left(
\begin{array}{c}
{\eta _{Lt}} \\
{\eta _{St}} \\
{\eta _{Ct}} \\
{\eta _{NIMt}}%
\end{array}%
{\normalsize \ }\right) ,
\]
where $L_{t},S_{t},$ and $C_{t}$ are level slope and curvature factors for the yield curve, respectively, and $\widehat{NIM}_{t}$ represent a measure of NIMs cleaned of observation error. The term $\mu _{x}$ denotes the mean of $x$. The terms ${\eta _{Lt}}$, ${\eta _{St}}$, ${\eta _{Ct}}$, ${\eta_{NIMt}}$ denote independently distributed innovations with joint covariance matrix $\Omega $. Our specification allows the yield curve factors to depend on their own lagged values but excludes lagged dependence on NIMs. By contrast, NIMs the specification allows for more general dependence of NIMs on lagged yield curve factors and and lagged NIMs.

The observation equations, which relate a set of $N$ yields (in our case, 12) and observed NIMs can be split into two groups. In the first group, equations for yields on treasuries of maturity $\tau $ take the form:
\[
r(\tau )=\left(
\begin{array}{cccc}
1 & \frac{1-e^{-\lambda \tau }}{\lambda \tau } & \frac{1-e^{-\lambda \tau }}{%
\lambda \tau }-e^{-\lambda \tau } & 0%
\end{array}%
\right) \left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}} \\
\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
\end{array}%
\right) +e(\tau )_{t},
\]%
where $e(\tau )_{t}$ is i.i.d. measurement error assumed to be uncorrelated
across observations. \bigskip In the second group, the equation for observed
NIMs takes the form%
\[
NIM_{t}=\left(
\begin{array}{cccc}
0 & 0 & 0 & 1%
\end{array}%
\right) \left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}} \\
\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
\end{array}%
\right) +e(NIM)_{t},
\]
where $e(NIM)_{t}$ is i.i.d. measurement error assumed to be uncorrelated across observations. Out-of-sample conditional forecasts for NIMs from this model, are obtained using the Kalman filter. We focus on the filtered estimates of NIMs given a path for all yields included in the model.

\noindent \textbf{3. Dynamic factor model with second-step regression for NIMs.} This model extracts level, slope, and curvature factors for the yield curve following Diebold and Li (2006). The factors are used to forecast NIMs in a separate second-step regression. The dynamic factor model we consider takes the form
\[
\left(
\begin{array}{c}
{L_{t+1}-\mu _{L}} \\
{S_{t+1}-\mu _{S}} \\
{C_{t+1}-\mu _{C}}%
\end{array}%
\right) {\normalsize \ =\ }\left(
\begin{array}{ccc}
a_{11} & a_{12} & a_{13} \\
a_{21} & a_{22} & a_{23} \\
a_{31} & a_{32} & a_{33}%
\end{array}%
\right) {\normalsize \ }\left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}}%
\end{array}%
\right) {\normalsize +}\left(
\begin{array}{c}
{\eta _{Lt}} \\
{\eta _{St}} \\
{\eta _{Ct}}%
\end{array}%
\right)
\]

\[
r(\tau )=\left(
\begin{array}{ccc}
1 & \frac{1-e^{-\lambda \tau }}{\lambda \tau } & \frac{1-e^{-\lambda \tau }}{%
\lambda \tau }-e^{-\lambda \tau }%
\end{array}%
\right) \left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}} 
\end{array}%
\right) +e(\tau )_{t},
\]
We obtain smoothed estimates of the factors $L_{t},S_{t},$ and $C_{t}$ conditional on information up until period $T.$ We estimate two alternative second-step forecasting equations:

\noindent \textbf{Alternative a. Multivariate} (all yield curve factors included in one NIM equation)
\[
NIM_{t}=c+\rho NIM_{t-1}+\gamma _{L}L_{t-1}+\gamma _{S}S_{t-1}+\gamma
_{C}C_{t-1}
\]
We use this equation to produce recursive forecasts for NIMs conditional on the the smoothed factors.

\noindent \textbf{Alternative b. Forecast combination} (each factor included in a separate equation and forecasts from each equation combined)
\[
NIM_{t}=c_{i}+\rho _{i}NIM_{t-1}+\gamma _{i}F_{i,t-1}+{\eta _{i,t},}
\]%
where $F_{i}\in \left\{ L,S,C\right\}.$ Recursive forecasts $NIM_{i,t+s/I}$ from each separate regression are then aggregated as:
\[
NIM_{t+s/I}=\sum_{i}\frac{NIM_{i,t+s/I}}{3},
\]

\textbf{4. Principal component regression}
Let $X$ a $T \times n$ matrix of $n$ Treasury yields.  Using the spectral decomposition, one obtains $X'X = VDV'$. Let $V_i$ be the normalized eigenvector corresponding to $i^{th}$ largest eigenvalue.
The $i^{th}$ principal component is defined as $PC_i = XV_i$. We estimate
\begin{equation*}
NIM_t = c + \rho NIM_{t-1} + \sum_{i=1}^3 \gamma_i PC_{i,t-1} + e_t.
\end{equation*}
using ordinary least squares and we construct recursive forecasts for net interest margins conditional on the principal components.

\noindent \textbf{5. Forecast combination with observed factors.} Define level factor $L$ to coincide with an average of the 3-month, 2-year and 10-year Treasury yields; define the slope factor as the difference between the 10-year and the 3-month yields; define the curvatures factor to be 2 times the 2-year yield minus the the sum of the 3-month and 10-year yields. Define $\left\{L,S,C \right\} ,$ the observed factors as in subsection 3.1 above.
\[
NIM_{t}=c_{i}+\rho _{i}NIM_{t-1}+\gamma _{i}O_{i,t-1}+{\eta _{i,t},}
\]%
where $O_{i}\in \left\{ L,S,C\right\} .$ Forecasts  from each separate regression, $NIM_{i,t+s/t},$are then aggregated as:

\noindent \textbf{6. VAR with observed factors.}  The VAR of order four includes the three observed factors defined as above and NIMs. Forecasts conditional on the factors are obtained using the Kalman filter. We take as conditional forecasts the filtered estimates of NIMs conditional on the pseudo-out-of-sample observations of the remaining variables in the VAR. This procedure efficiently considers the  covariance between reduced-form residuals in the VAR.

\noindent \textbf{7. No-change forecast.} Taking the forecasting model to be a random walk without a drift implies a no change forecast at all horizons $s$: $NIM_{t+s/t}=NIM_{t}${\normalsize . }

\section{Data}

The baseline sample starts in 1989q1 and stops in 2008q2. Our dataset would allow us to start even earlier, in 1985q4, but we delay the start of the estimation sample for two reasons.  First, there is a peculiar spike evident in aggregated NIMs in Figure \ref{figure_nims_rates} in 1988q4.  This large spike stems from late payments from Brazil during the Latin American debt crisis. The end point of 2008q2 is dictated by the attainment of the zero lower bound on nominal interest rates.\footnote{To estimate the dynamic factor models described under 2. and 3., we use observations between 1985q4 and 1989q4 as a training sample.}

\subsection{Net interest margins}

\vspace{-0.1in} The data on bank net interest margins (NIMs) that we use is from the quarterly Consolidated Reports of Condition and Income (Call Report) that every national, state member, and insured nonmember bank is required to file on the last day of each quarter by the Federal Financial Institutions Examination Council (FFIEC). The Federal Deposit Insurance Corporation is tasked as the overseer, collecting and reviewing all submissions. Call Report data used in this analysis are cleaned and adjusted for bank mergers and acquisitions, using structure data from the National Information Clearinghouse (NIC) on mergers and acquisitions. Foreign entities are excluded and domestic subsidiaries are aggregated up to the parent, bank-holding-company (BHC), level.\footnote{NIM data are adjusted for mergers between commercial banks by comparing balance sheet values of interest income, interest expenses, and interest-earning assets at the end of the quarter with those at the beginning of the quarter, accounting for amounts acquired or lost during the period because of mergers -- see the appendix in English and Nelson (1998).}

NIMs in our models are expressed as the annualized percentage of net interest income over interest-earning assets. To get an aggregate banking sector measure of NIMs, we aggregate NIMs for the top 25 BHCs, as ranked by total assets, which is assessed quarterly.

\subsection{Treasury yields}

The yields data are derived using a smoothing technique from Gurkaynak, et al. (2007), based on Nelson and Siegel (1987) and Svensson (1994), which allows for daily measures of an off-the-run Treasury yield curve. We use quarterly yields for twelve maturities in our models: 3 month, 6 month, 9 month, 1 year, 2 year, 3 year, 5 year, 7 year, 10 year, 15 year, 20 year, and 30 year.\footnote{Daily yields are published with a two-day lag under the base mnemonic SVENY at  www.federalreserve.gov/econresdata/researchdata/feds200628\_1.html.} The yields included are quarterly averages of daily yields.

\subsection{Micro-banking variables}

\vspace{-0.1in} The main variable that we use to capture the degree of competition that banks face is the size of the traditional
banking industry (as measured by the assets in that industry) relative to the size of the total banking industry, that is, both
traditional and shadow banking (as measured by the combined assets of these these two industries). We use this measure for the
reason that competition from the shadow banking sector is the most relevant form of increased competition faced by the
traditional banking sector.  Importantly, competition from the shadow banking sector affects both sides of traditional banks'
balance sheets. For example, it puts upward pressure on the deposit rate that traditional banks must pay to attract deposits---since the shadow banking sector offers alternatives to deposits like money market mutual funds---and it puts downward pressure on the lending rate that traditional banks can earn---since the shadow banking sector offers alternatives to bank loans like loans from finance companies. This measure can be easily calculated from the U.S. Financial Accounts (previously known as the Flow of Funds Accounts), where the traditional banking sector consists of commercial banks, savings institutions, and credit unions and the shadow banking sector consists of broker-dealers, ABS issuers, finance companies, mortgage pools, and funding corporations.

In papers that have modeled NIMs as a function of more microeconomic variables interest rate volatility has been measured by
the quarterly variance of weekly Treasury interest rate series. (See, for example, Ho and Saunders, 1981, and Saunders
and Schumacher, 2000.)  Typically, also, it is the quarterly variance of several Treasury interest rate series that are
considered---such as, the 3-month rate, 1-year rate, 2-year rate, 3-year rate, and 5-year rate---with the best fitting series then
included in the regression.  We also consider the Merrill Lynch MOVE index, which is the bond market equivalent to the VIX.

\section{Results}

To compare the performance of the models, we consider pseudo-out-of-sample forecasts. We use a rolling window for estimation. The shortest window spans 1989q4 to 1999q4. The assessment window spans 2000q1 to 2008q3. We focus on forcasts from 1 to 10 steps ahead. In constructing the out-of-sample forecasts, we condition on the observation of the right-hand-side variables in each model, with the exception of lagged NIMs.

Table \ref{table_rmse_baseline} reports root mean square errors (RMSE) for the pseudo-out-of-sample forecasts.  Model 1., the equal weighted forecast combination, is the only model to have lower RMSE's relative to the no-change forecasts for Model 6. In relative terms, the advantage of Model 1. over Model 6. increases as the forecast horizon lengthens. Model 2, performs uniformly worse than model 6. but, excluding Model 1, it performs better than any other model we considered. Models 3a. and 3b. do not incorporate the information in the variance/covariance matrix of the shock in the state equation of Model 2.  The choice of variables for the forecast combination in model 4. is not as broad as for 1. Finally, the proliferation of parameters in the VAR of model 5. is associated with the worst performance, overall.

Table \ref{table_rmse_competition} considers variants of the model that include the share of assets of the shadow banking sector. \footnote{In this case, we exclude Model 2. from the comparison.}  In no model, does that addition  improve the performance of the conditional forecasts.

Table \ref{table_rmse_combination} considers the performance of alternative variants of Model 1. Using either our ``observed factors,'' or factors derived as smoothed estimates from the yield curve model set up as part of Model 3., does not improve the forecast performance.  By contrast, pruning out some of the yields, lowers the RMSEs for the forecasts even further. A forecast combination approach that includes models conditional on the 3-month Treasury yields and the 10-year treasury yields has the best out-of-sample performance across the models we considered.

Figure \ref{figure_scenario} uses the forecast combination model that includes the 3-month and the 10-year yields only to forecast NIMs conditional on the three CCAR 2013 supervisory scenarios.  The severely adverse scenario implies a downward shift of the profile of the yield curve accompanied by a sharp flattening. The severe scenario implies an upward shift of the yield curve with little change in the slope. In terms of changes in the yield curve, these two scenarios capture vastly different configurations.  The implications of the point forecasts appear sensible. The severely adverse scenario reduces NIMs, but it takes a few quarters for the implications of the change in yields to transmit to NIMs. By contrast, the upward parallel shift in the yield curve is predicted to boost NIMs.

The figure, provides a visual interpretation of the RMSEs considered thus far. The bands surrounding the central forecast are constructed adding and subtracting to the point forecast the RMSEs for the forecast combination model. Strikingly, the size of the uncertainty bands dwarfs the variation implied by the scenarios for NIMs. The two severe scenarios are statistically indistinguishable from the baseline and from each other.

\section{Conclusion}

The unweighted forecast combination model for NIMs based on just two yields outperforms the other forecast models we considered. No other model comes close to outperforming a no-change forecast.  This is only a small victory as the variance of the forecast error is large even when compared to the observed variation of NIMs. Accordingly, even scenarios that have vastly different implications for the configuration of interest rates contain relatively little information to forecast NIMs.

Our findings have implications beyond the study of net interest margins. Scenario-based stress tests have been used across the world to rebuild confidence in the banking sector in periods of crises. Yet, they are predicated on the assumption that different scenarios for the evolution of the macroeconomy have sufficiently different implications for the evolution of key balance sheet items.  Yet, an array of time series methods found by others to be near the frontier of efficiency produces projections conditional on radically different scenarios for yields that statistically indistinguishable.

Supervisory authorities that perform stress tests certainly have access to a wealth of supervisory information beyond what  the publicly available data that we have used. Yet, if the conclusions of stress tests are mostly predicated on supervisory sources, as our study implies, communication with the public is even more challenging. In that case, the effects of confidence of stress tests rely squarely on the reputation of the supervising authority, one which might be tarnished exactly in a period of crisis.

\newpage
{\normalsize \clearpage
\begin{sidewaystable}
\center
\caption{RMSE for baseline specifications} \label{table_rmse_baseline}
\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
\hline
&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\
\hline
1. Simple Forecast Combination             &0.107&0.133&0.148&0.168&0.187&0.202&0.213&0.209&0.216&0.223\\
2. DFM                                                       &0.122&0.148&0.169&0.212&0.247&0.278&0.316&0.340&0.373&0.404\\
3a. DFM with 2nd Step Regression           &0.118&0.167&0.209&0.260&0.308&0.356&0.398&0.433&0.474&0.511\\
3b. DFM with Forecast Combination          &0.112&0.159&0.199&0.247&0.300&0.351&0.399&0.438&0.489&0.535\\
4. Forecast Combination of Observed Factors&0.118&0.169&0.216&0.273&0.336&0.399&0.456&0.509&0.567&0.622\\
5. VAR on Observed Factors                 &0.166&0.209&0.263&0.302&0.342&0.375&0.380&0.430&0.451&0.499\\
6. No-Change Forecast                      &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
\hline
\end{tabular}
\end{sidewaystable}
}

\newpage
{\normalsize \clearpage
\begin{sidewaystable}
\caption{RMSEs for models that include the share of asset of the shadow banking sector. Estimation sample: 1989Q2-2008Q2} \label{table_rmse_competition}
\center
\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
\hline
&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\
\hline
1. Simple Forecast Combination             &0.110&0.143&0.165&0.192&0.222&0.248&0.267&0.274&0.287&0.300\\
3a. DFM with 2nd Step Regression           &0.121&0.170&0.212&0.262&0.312&0.360&0.403&0.436&0.478&0.515\\
3b. DFM with Forecast Combination          &0.106&0.140&0.165&0.193&0.222&0.248&0.267&0.274&0.287&0.300\\
4. Forecast Combination of Observed Factors&0.109&0.141&0.165&0.193&0.222&0.249&0.267&0.274&0.288&0.300\\
5. VAR on Observed Factors                 &0.344&0.396&0.476&0.494&0.521&0.569&0.631&0.670&0.699&0.718\\
6. No-Change Forecast                      &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
\hline
\end{tabular}
\end{sidewaystable}
}

\newpage \clearpage
\begin{sidewaystable}
\caption{RMSEs for variants of the simple forecast combination model} \label{table_rmse_combination}
\center
\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
\hline
&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\
\hline
All yields              &0.107&0.133&0.148&0.168&0.187&0.202&0.213&0.209&0.216&0.223\\
Observed Factors        &0.118&0.169&0.216&0.273&0.336&0.399&0.456&0.509&0.567&0.622\\
Smoothed Factors        &0.112&0.159&0.199&0.247&0.300&0.351&0.399&0.438&0.489&0.535\\
3-month, 2-year, 10-year&0.105&0.130&0.141&0.160&0.178&0.197&0.202&0.198&0.209&0.217\\
3-month, 10-year        &0.105&0.129&0.144&0.159&0.179&0.196&0.200&0.200&0.209&0.217\\
\hline
\end{tabular}
\end{sidewaystable}

\newpage \clearpage
\newpage \clearpage
\begin{figure}
\caption{Interest Income, Interest Expenses and Short-term Treasury Rates} \label{figure_nims_components}
\center
\includegraphics[scale=0.85]{figure_nims_components.ps}
\end{figure}

\begin{figure}
%[tbp]
\caption{NIMs and Treasury Yields} \label{figure_nims_rates}
\center
\includegraphics[scale=0.85]{figure_nims_rates.ps}
\end{figure}

\newpage \clearpage
\begin{figure}
%[tbp]
\caption{NIMs and Observed Yield Factors} \label{figure_nims_factors}
\center
\includegraphics[scale=0.85]{figure_nims_factors.ps}
\end{figure}

\clearpage
\begin{figure}
\flushleft
\caption{NIMs and Competition -- The Asset Share of the Shadow Banking Sector} \label{figure_nims_competition}

\hspace{-2.5cm} \includegraphics[scale=0.85]{figure_nims_competition.ps}
\end{figure}

\clearpage
\begin{figure}
\flushleft
\caption{Forecast for NIMs Conditional on CCAR 2012 Scenarios} \label{figure_scenario}

 \includegraphics[scale=0.85]{figure_scenario.ps}
\end{figure}

\end{document}